Puck Predictions: Unraveling the NHL Game Forecasting Riddle

### 2024-03-29 Latex/draft.tex
Pushing latex changes.

\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,amsfonts,hyperref, enumitem, tikz}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

\subsection{Regression and XGBoost}

@@

To simulate hockey games, we created a Markov Chain where the states are a tuple of three consecutive
events that occurred. For example, if the home team won a faceoff, the home team lost the puck, and 
then the away team shot the puck and missed, the markov chain would look like figure \ref{fig:markov_chain_sample}.

>>>>

\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,amsfonts,hyperref, enumitem, tikz}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{subfigure}

@@

\subsection{XGBoost}

In our NHL analysis research, we partitioned the dataset into segments 
corresponding to individual teams' games over multiple seasons. 
This enabled the creation of time series data in the form of a state vector, capturing the 
play-by-play dynamics of each team's matches. We then trained 
separate XGBoost models for each team to learn their unique patterns 
and strategies. Leveraging the {\tt .predict\_proba()} method of XGBoost, we generated 
probabilistic predictions at various stages of a game, allowing us to 
plot the evolving probability of each team winning throughout the match
(see~\ref{fig:xgboost_res}). 
This approach provided insights into momentum shifts (when one team 
has an advantage due to them having more players on the ice due to 
a penalty by the opposing team), key moments (such as goals and 
penalties, and critical plays influencing game outcomes).

% Images
\begin{figure}[H]
    \centering
    \subfigure[]{\includegraphics[width=0.9\textwidth]{images/good_xgb_example.png}} 
    \subfigure[]{\includegraphics[width=0.9\textwidth]{images/good_xgb_example(2).png}} 
    \caption{For both of these plots, the home team is above the the 50\% line, with the
    away team below. The title has the final score of the game. \\
        \textbf{(a)} The live win probabilities for the Nashville Predators at the Vegas Golden Knights. 
    Notice the probability shifts after one team scores a goal.  \\
        \textbf{(b)} The live win probabilities for the Calgary Flames at the Winnipeg Jets.
        Notice the stark changes in probabilities after each goal. Also notice the
        shift in probabilities even when no goals are scored. This suggests
        the XGBoost model is able to pick up on momentum shifts.}
    \label{fig:xgboost_res}
\end{figure}

@@

To simulate hockey games, we created a Markov Chain where the states are a tuple of three consecutive
events that occurred. For example, if the home team won a faceoff, the home team lost the puck, and 
then the away team shot the puck and missed, the markov chain would look like figure~\ref{fig:markov_chain_sample}.
